**Machine learning identifies unique taxa differentiating proximal and distal

### 2017-06-14 submission/manuscript.Rmd
more code in manuscript

alltyc[10] <- as.factor(alltyc[10])

@@

>>>>

@@

```{r randomforest}
#create dataframe with relative abundances
rel_abund <- 100*shared_file/unique(apply(shared_file, 1, sum))

#Create vector of OTUs with median abundances >1%
OTUs_1 <- apply(rel_abund, 2, max) > 1
OTU_list <- colnames(rel_abund)[OTUs_1]
#get df of just top OTUs
rel_abund_top <- rel_abund[, OTUs_1]
rel_meta <- merge(meta_file, rel_abund_top, by.x='group', by.y="row.names")

seed <- 1
n_trees <- 2001

source('../code/random_functions.R')

#####RandomForest###########################################################################################
#build randomForest model for each location comparison using randomize_loc function 
rf_left <- randomize_loc(rel_meta, "LB", "LS") #OOB 10.26%
rf_right <- randomize_loc(rel_meta, "RB", "RS") #OOB 53%
rf_bowel <- randomize_loc(rel_meta, "LB", "RB") #OOB 25.64%
rf_lumen <- randomize_loc(rel_meta, "LS", "RS") #OOB 69.23%

#and for each site
rf_all <- randomize_site(rel_meta, "mucosa", "stool")

#####AUCRF#####################################################################################

# create RF model with AUCRF outputs top OTUs
aucrf_data_left_bs <- auc_loc(rel_meta, "LB", "LS")
aucrf_data_LRbowel <- auc_loc(rel_meta, "LB", "RB")
aucrf_data_right_bs <- auc_loc(rel_meta, "RB", "RS")
aucrf_data_LRlumen <- auc_loc(rel_meta, "LS", "RS")
aucrf_data_allum <- auc_site(rel_meta, "mucosa", "stool")

########Cross-Validation#################################################################
#10 fold cross validation for all lumen vs mucosa 
iters <- 100
cv10f_aucs <- c()
cv10f_all_resp <- c()
cv10f_all_pred <- c()
for(j in 1:iters){
  set.seed(j)
  sampling <- sample(1:nrow(aucrf_data_allum),nrow(aucrf_data_allum),replace=F)
  cv10f_probs <- rep(NA,78)
  for(i in seq(1,77,7)){
    train <- aucrf_data_allum[sampling[-(i:(i+6))],]
    test <- aucrf_data_allum[sampling[i:(i+6)],]
    set.seed(seed)
    temp_model <- AUCRF(site~., data=train, pdel=0.99, ntree=500)
    cv10f_probs[sampling[i:(i+6)]] <- predict(temp_model$RFopt, test, type='prob')[,2]
  }
  cv10f_roc <- roc(aucrf_data_allum$site~cv10f_probs)
  cv10f_all_pred <- c(cv10f_all_pred, cv10f_probs)
  cv10f_all_resp <- c(cv10f_all_resp, aucrf_data_allum$site)
  cv10f_aucs[j] <- cv10f_roc$auc #stores aucs for all iterations, can use to calc IQR
}
cv10f_roc <- roc(cv10f_all_resp~cv10f_all_pred)

#10fold CV for L lumen vs L mucosa
iters <- 100
cv10f_aucs <- c()
cv10f_all_resp_left_bs <- c()
cv10f_all_pred_left_bs <- c()
for(j in 1:iters){
  set.seed(j)
  sampling <- sample(1:nrow(aucrf_data_left_bs),nrow(aucrf_data_left_bs),replace=F)
  cv10f_probs <- rep(NA,39)
  for(i in seq(1,36,4)){
    train_left_bs <- aucrf_data_left_bs[sampling[-(i:(i+3))],]
    test_left_bs <- aucrf_data_left_bs[sampling[i:(i+3)],]
    set.seed(seed)
    temp_model_left_bs <- AUCRF(location~., data=train_left_bs, pdel=0.99, ntree=500)
    cv10f_probs[sampling[i:(i+3)]] <- predict(temp_model_left_bs$RFopt, test_left_bs, type='prob')[,2]
  }
  cv10f_roc_left_bs <- roc(aucrf_data_left_bs$location~cv10f_probs)
  cv10f_all_pred_left_bs <- c(cv10f_all_pred_left_bs, cv10f_probs)
  cv10f_all_resp_left_bs <- c(cv10f_all_resp_left_bs, aucrf_data_left_bs$location)
  cv10f_aucs[j] <- cv10f_roc_left_bs$auc #stores aucs for all iterations, can use to calc IQR
}
cv10f_roc_left_bs <- roc(cv10f_all_resp_left_bs~cv10f_all_pred_left_bs)

#10fold CV for R lumen vs R mucosa
iters <- 100
cv10f_aucs <- c()
cv10f_all_resp_right_bs <- c()
cv10f_all_pred_right_bs <- c()
for(j in 1:iters){
  set.seed(j)
  sampling <- sample(1:nrow(aucrf_data_right_bs),nrow(aucrf_data_right_bs),replace=F)
  cv10f_probs <- rep(NA,39)
  for(i in seq(1,36,4)){
    train_right_bs <- aucrf_data_right_bs[sampling[-(i:(i+3))],]
    test_right_bs <- aucrf_data_right_bs[sampling[i:(i+3)],]
    set.seed(seed)
    temp_model_right_bs <- AUCRF(location~., data=train_right_bs, pdel=0.99, ntree=500)
    cv10f_probs[sampling[i:(i+3)]] <- predict(temp_model_right_bs$RFopt, test_right_bs, type='prob')[,2]
  }
  cv10f_roc_right_bs <- roc(aucrf_data_right_bs$location~cv10f_probs)
  cv10f_all_pred_right_bs <- c(cv10f_all_pred_right_bs, cv10f_probs)
  cv10f_all_resp_right_bs <- c(cv10f_all_resp_right_bs, aucrf_data_right_bs$location)
  cv10f_aucs[j] <- cv10f_roc_right_bs$auc #stores aucs for all iterations, can use to calc IQR
}
cv10f_roc_right_bs <- roc(cv10f_all_resp_right_bs~cv10f_all_pred_right_bs)

#10 fold cross validation for L vs R mucosa

iters <- 100
cv10f_aucs_muc <- c()
cv10f_all_resp_muc <- c()
cv10f_all_pred_muc <- c()
for(j in 1:iters){
  set.seed(j)
  sampling_muc <- sample(1:nrow(aucrf_data_LRbowel),nrow(aucrf_data_LRbowel),replace=F)
  cv10f_probs_muc <- rep(NA,39)
  for(i in seq(1,36,4)){
    train_muc <- aucrf_data_LRbowel[sampling_muc[-(i:(i+3))],]
    test_muc <- aucrf_data_LRbowel[sampling_muc[i:(i+3)],]
    set.seed(seed)
    temp_model_muc <- AUCRF(location~., data=train_muc, pdel=0.99, ntree=500)
    cv10f_probs_muc[sampling_muc[i:(i+3)]] <- predict(temp_model_muc$RFopt, test_muc, type='prob')[,2]
  }
  cv10f_roc_muc <- roc(aucrf_data_LRbowel$location~cv10f_probs_muc)
  cv10f_all_pred_muc <- c(cv10f_all_pred_muc, cv10f_probs_muc)
  cv10f_all_resp_muc <- c(cv10f_all_resp_muc, aucrf_data_LRbowel$location)
  cv10f_aucs_muc[j] <- cv10f_roc_muc$auc #stores aucs for all iterations, can use to calc IQR
}
cv10f_roc_muc <- roc(cv10f_all_resp_muc~cv10f_all_pred_muc)

#10 fold cross validation for L vs R lumen
iters <- 100
cv10f_aucs_lum <- c()
cv10f_all_resp_lum <- c()
cv10f_all_pred_lum <- c()
for(j in 1:iters){
  set.seed(j)
  sampling_lum <- sample(1:nrow(aucrf_data_LRlumen),nrow(aucrf_data_LRlumen),replace=F)
  cv10f_probs_lum <- rep(NA,39)
  for(i in seq(1,36,4)){
    train_lum <- aucrf_data_LRlumen[sampling_lum[-(i:(i+3))],]
    test_lum <- aucrf_data_LRlumen[sampling_lum[i:(i+3)],]
    set.seed(seed)
    temp_model_lum <- AUCRF(location~., data=train_lum, pdel=0.99, ntree=500)
    cv10f_probs_lum[sampling_lum[i:(i+3)]] <- predict(temp_model_lum$RFopt, test_lum, type='prob')[,2]
  }
  cv10f_roc_lum <- roc(aucrf_data_LRlumen$location~cv10f_probs_lum)
  cv10f_all_pred_lum <- c(cv10f_all_pred_lum, cv10f_probs_lum)
  cv10f_all_resp_lum <- c(cv10f_all_resp_lum, aucrf_data_LRlumen$location)
  cv10f_aucs_lum[j] <- cv10f_roc_lum$auc #stores aucs for all iterations, can use to calc IQR
}
cv10f_roc_lum <- roc(cv10f_all_resp_lum~cv10f_all_pred_lum)

```{r important_otus}
